Rusty Linux: Advances in Rust for Linux Kernel Development

### 2024-06-05 paper.tex
Update on Overleaf.

To begin our study, we must first confirm the need for a SLR. The recent report released by
ONCD~\cite{United_States_Gov2024-pp} has conveniently done this job for us by compiling a report
detailing the need for research in the domain of memory safety. While the ONCD report detailed a two-pronged approach, for this reflection paper, we will be focusing on a memory-safe programming language,
specifically Rust. While there are many modern memory-safe programming languages available for
software developers to use, Rust is one of the few languages that is feasible to use when developing
operating system kernels due to its lack of runtime and garbage collector. Even the Linux kernel is
in the early stages of adding real support for
Rust~\cite{The_kernel_development_community_undated-iw} thus, we will focus on Rust as a primary
candidate to replace the aging C programming language.

@@

The development of operating system kernels in Rust introduces several unique challenges and
limitations compared to more traditional languages like C. These challenges arise from Rust’s strict
safety guarantees, its relatively recent adoption in systems programming, and the inherent
complexities of kernel development. This section discusses these challenges, limitations, and the
lessons learned from various projects and research efforts. We summarize our findings in
table~\ref{tab:RQ3}.

@@

\subsection{RQ4: Lessons Learned}
Klimt et al.~\cite{Klimt2023-ob} details the lessons learned and challenges when implementing
Theseus. First they found that it is impossible to write a complete operating system in 100\% safe
Rust. For example, when writing a memory management system, raw pointers must be used to
modify the hardware. Despite the limitations imposed by the hardware, the authors describe how
Theseus leveraged intralingual design to maximize the compiler's role in enforcing correctness. By
leveraging Rust's type system and borrow checker memory safety, correct ownership transfer can be
achieved at a higher level than what could be done in C. The authors also detailed some of the
limitations of intralingual design, such as not being as expressive as many other formal verification
techniques due to the limited invariance that the type system can enforce. One of the most
important lessons learned was the insight that a linear type system itself cannot guarantee
the uniqueness of the resource represented, such as when a memory resource may overlap. The authors
introduce the idea of using a hybrid approach of verification where they leverage both the linear
type system and an SMT solver. The authors also explore the bootstrapping problem with rust
systems. How do we ensure that ownership of memory is sound in the operating system itself? In
operating systems written in C, the kernel provides the ownership root to applications running on
top. Finding a new ownership root is an open research question with regard to using Rust for
operating system development.

@@

Definition Ambiguity: The term "Kernel development in Rust" encompasses a wide range of activities,
from minor module development to complete subsystem rewrites. Variations in how researchers and
developers interpret this integration could lead to inconsistent findings. Selection Criteria: The
inclusion and exclusion criteria for selecting relevant studies might inadvertently bias the review
towards certain types of research, such as those reporting successful integration while
underrepresenting studies detailing challenges and failures. Compared to C, Rust is a brand new
language that is still finding its way in the world.

@@

\begin{itemize}
  \item Publication Bias: Studies demonstrating positive outcomes of using Rust for kernel
    development may be more likely to be published and cited, leading to an overestimation of its
    benefits.
  \item Quality of Studies: The methodological rigor of the included studies varies, and some may
    suffer from design flaws or biases that were not adequately controlled for, thus affecting the
    reliability of their findings.
\end{itemize}

@@

\begin{itemize}
  \item Context-Specific Findings: Many studies focus on specific kernel modules or use-cases, which
    may not be representative of the broader kernel environment. The success of Rust in isolated
    components does not necessarily translate to the entire kernel.
  \item Temporal Changes: The rapidly evolving nature of both the Rust programming language and the
    Linux kernel means that findings from older studies may no longer be applicable, as improvements
    and changes in both domains can alter the landscape significantly.
\end{itemize}

>>>>

To begin our study, we must first confirm the need for a SLR. The recent report released by
ONCD~\cite{United_States_Gov2024-pp} has conveniently done this job for us by compiling a report
detailing the need for research in the domain of memory safety. While the ONCD report detailed a two-pronged approach, for this reflection paper, we will be focusing on a memory-safe programming language,
specifically Rust. While there are many modern memory-safe programming languages available for
software developers to use, Rust is one of the few languages that is feasible to use when developing
operating system kernels due to its lack of runtime and garbage collector~\cite{Ojeda2021-ul}. Even the Linux kernel is
in the early stages of adding real support for
Rust~\cite{The_kernel_development_community_undated-iw} thus, we will focus on Rust as a primary
candidate to replace the aging C programming language.

@@

The development of operating system kernels in Rust introduces several unique challenges and
limitations compared to more traditional languages like C. These challenges arise from Rust’s strict
safety guarantees, its relatively recent adoption in systems programming, and the inherent
complexities of kernel development~\cite{Ojeda2021-ul}. This section discusses these challenges, limitations, and the
lessons learned from various projects and research efforts. We summarize our findings in
table~\ref{tab:RQ3}.

@@

\subsection{RQ4: Lessons Learned}

Klimt et al.~\cite{Klimt2023-ob} details the lessons learned and challenges when implementing
Theseus. First they found that it is impossible to write a complete operating system in 100\% safe
Rust. For example, when writing a memory management system, raw pointers must be used to
modify the hardware. Despite the limitations imposed by the hardware, the authors describe how
Theseus leveraged intralingual design to maximize the compiler's role in enforcing correctness. By
leveraging Rust's type system and borrow checker memory safety, correct ownership transfer can be
achieved at a higher level than what could be done in C. The authors also detailed some of the
limitations of intralingual design, such as not being as expressive as many other formal verification
techniques due to the limited invariance that the type system can enforce. One of the most
important lessons learned was the insight that a linear type system itself cannot guarantee
the uniqueness of the resource represented, such as when a memory resource may overlap. The authors
introduce the idea of using a hybrid approach of verification where they leverage both the linear
type system and an SMT solver. The authors also explore the bootstrapping problem with rust
systems. In operating systems written in C, the kernel provides the ownership root to applications running on
top. Finding a new ownership root is an open research question with regard to using Rust for
operating system development.

@@

@@

\begin{itemize}
  \item Publication Bias: Studies demonstrating positive outcomes of using Rust for kernel
    development may be more likely to be published and cited, leading to an overestimation of its
    benefits.
  \item Quality of Studies: The methodological rigor of the included studies varies, and some may
    suffer from design flaws or biases that were not adequately controlled for, thus affecting the
    reliability of their findings.
  \item  Selection Criteria: The inclusion and exclusion criteria for selecting relevant studies might inadvertently bias the review         towards certain types of research, such as those reporting successful integration while
    underrepresenting studies detailing challenges and failures.
\end{itemize}

@@

\begin{itemize}
  \item Context-Specific Findings: Many studies focus on specific kernel modules or use-cases, which
    may not be representative of the broader kernel environment. The success of Rust in isolated
    components does not necessarily translate to the entire kernel.
  \item Temporal Changes: The rapidly evolving nature of both the Rust programming language and the
    Linux kernel means that findings from older studies may no longer be applicable, as improvements
    and changes in both domains can alter the landscape significantly.
    \item Definition Ambiguity: The term "Kernel development in Rust" encompasses a wide range of activities,
    from minor module development to complete subsystem rewrites. Variations in how researchers and
    developers interpret this integration could lead to inconsistent findings.
\end{itemize}
